---
title: "ANLY 699 Missing Data and Outliers Analysis"
author: "Yizhou Liu"
date: "09/05/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Preparation

## Import Data
```{r import data}
dt <-  read.csv("/Users/yizhouliu/Documents/HU/ANLY 699/Social_Anxiety_Survey_Master.csv")
```

## Remove NAs (or impute). 10pts
```{r check missing values}
missingcells = sum(is.na(dt))
print(missingcells)

# No missing data in this dataset.
```

```{r data pre-processing}
# remove the first column "Timestamp" as we don't need it in this study
dt <- subset(dt, select = -Timestamp)
```
```{r convert data types}
# convert data type of "Gender"
sexType <- c(Male = 1, Female = 2, Other = 3)
dt$Gender <- sexType[dt$Gender]

# convert data type of "Student"
isStudent <- c(Yes = 1, No = 0, Other = 3)
dt$Student <- isStudent[dt$Student]

# convert data type of "Age"
ageGroup <- c("20 or below" = 1, "21 - 30" = 2, "31 - 40" = 3 )
dt$Age <- ageGroup[dt$Age]

# convert data type of "Marital"
maritalStatus <- c(Single = 1, Married = 2)
dt$Marital <- maritalStatus[dt$Marital]

# convert data type of "Q1"
smUseTime <- c("Less than 1 hours per day" = 1, "Less than 2 hours per day" = 2, "Less than 3 hours per day" = 3, "Less than 4 hours per day" = 4, "4 hours and above per day" = 5)
dt$Q1 <- smUseTime[dt$Q1]


```
```{r}
lapply(dt, as.numeric)
```



# Data Normalization
```{r data normalization}
#Create a function to normalize variables
normalize_data <- function(val) { return ((val - min(val)) / (max(val) - min(val))) }

```
```{r}
norm_data <- as.data.frame(apply(dt, 2, normalize_data))
summary(norm_data)
```


## Create a subset without dependent variable (outcome/response). 10pts
```{r Create a subset without dependent variable (outcome/response)}
# exclude columns Q6 and Q7
subdt <- norm_data[c(-10, -11)]
subdt
```
# K-Means

## Find the optimal number of clusters (elbow, gap or silhouette methods). Plot the Figure. 10pts
```{r Find the optimal number of clusters}
#Elbow Method for finding the optimal number of clusters
set.seed(123)
k.max <- 15
data <- subdt
library(factoextra)


fviz_nbclust(data, kmeans, method = "wss", k.max = 24) + theme_minimal() + ggtitle("the Elbow Method")
```
##  Explain how you decided on the optimal number. 10pts
- Therefore, for k=4 the between_ss/total_ss ration tends to change slowly and remain less changing as compared to other ks. So data k=4 should be a good choice for number of cluesters, but k=5 is also a good candidate.

## Perform the K-Means cluster analysis. Do not print results. 10pts
```{r K-Means cluster analysis}
library(stats)
set.seed(123)
km.res <- kmeans(data, 4, nstart = 25)
```

## Visualize cluster. Consider adding Y labels (see reading and practice for code). 10pts
```{r adding Y labels}
set.seed(123)
k.max <- 15
data2 <- norm_data
fviz_nbclust(data2, kmeans, method = "wss", k.max = 24) + theme_minimal() + ggtitle("the Elbow Method")
```
## Provide your interpretation of the cluster visualization. Does it provide you with additional insights for your data? 10p
- After adding Y lables, for k=5 the between_ss/total_ss ration tends to change slowly and remain less changing as compared to other ks.


# Hierarchical Cluster

##Build hierarchical cluster. Explain what linkage is used (you can try several linkage methods). 10pts
```{r Build hierarchical cluster}
clusters <- hclust(dist(norm_data))
```

- By default, here the complete linkage was used.


## Plot hierarchical cluster. 10pts
```{r Plot hierarchical cluster}
plot(clusters)
```
## How many clusters do you observe. Interpret the results. Does it provide you with additional insights? 10pt (Make sure to use a proper academic writing style to describe results - see the required book on academic writing)

- From the plot, we can see there are 2 clusters.





